\chapter{Appendix: Interpersonal Reactivity Index}
\label{chap_app_iri}
Adapted: from Davis, Mark H ``Measuring individual differences in empathy: Evidence for a multidimensional approach.'' In: \textit{Journal of personality and social psychology} (1983) p. 113.

\section{Instruction}
The following statements inquire about your thoughts and feelings in a variety of situations. For  each item, indicate how well it describes you by choosing the appropriate number on the scale: 1, 2, 3, 4, or 5. (1 - Does not describe me well, 5 - Describes me very well)

READ EACH ITEM CAREFULLY BEFORE RESPONDING. Answer as honestly as you can.

\section{Questionnaire}
\begin{enumerate}
\item I daydream and fantasize, with some regularity, about things that might happen to me.
(FS)
\item I often have tender, concerned feelings for people less fortunate than me. (EC)
\item I sometimes find it difficult to see things from the ``other guy's'' point of view. (PT) (-)
\item Sometimes I don't feel very sorry for other people when they are having problems. (EC)
(-)
\item I really get involved with the feelings of the characters in a novel. (FS)
\item In emergency situations, I feel apprehensive and ill-at-ease. (PD)
\item I am usually objective when I watch a movie or play, and I don't often get completely
caught up in it. (FS) (-)
\item I try to look at everybody's side of a disagreement before I make a decision. (PT)
\item When I see someone being taken advantage of, I feel kind of protective towards them.
(EC)
\item I sometimes feel helpless when I am in the middle of a very emotional situation. (PD)
\item I sometimes try to understand my friends better by imagining how things look from
their perspective. (PT) 
Self Report Measures for Love and Compassion Research: Empathy
\item Becoming extremely involved in a good book or movie is somewhat rare for me. (FS) (-)
\item When I see someone get hurt, I tend to remain calm. (PD) (-)
\item Other people's misfortunes do not usually disturb me a great deal. (EC) (-)
\item If I'm sure I'm right about something, I don't waste much time listening to other
people's arguments. (PT) (-)
\item After seeing a play or movie, I have felt as though I were one of the characters. (FS)
\item Being in a tense emotional situation scares me. (PD)
\item When I see someone being treated unfairly, I sometimes don't feel very much pity for
them. (EC) (-)
\item I am usually pretty effective in dealing with emergencies. (PD) (-)
\item I am often quite touched by things that I see happen. (EC)
\item I believe that there are two sides to every question and try to look at them both. (PT)
\item I would describe myself as a pretty soft-hearted person. (EC)
\item When I watch a good movie, I can very easily put myself in the place of a leading
character. (FS)
\item I tend to lose control during emergencies. (PD)
\item When I'm upset at someone, I usually try to ``put myself in his shoes'' for a while. (PT)
\item When I am reading an interesting story or novel, I imagine how I would feel if the
events in the story were happening to me. (FS)
\item When I see someone who badly needs help in an emergency, I go to pieces. (PD)
\item Before criticizing somebody, I try to imagine how I would feel if I were in their place.
(PT) 
\end{enumerate}